The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-10-29 theoretical-consensus/article/article.tex
add page 06

\usepackage[pdftex]{graphicx}
\graphicspath{ {./images/} }

@@

The advancements of the internet 
more than potentiated the revolution and what we now call the permissionless consensus problem was 
finally born. Without forgetting the previous attempts, the first practical permissionless consensus algorithm was 
proposed by Nakamoto in \cite{nakamoto2008bitcoin}. It is a proof-of-work consensus protocol that resembles a "replicated 
state machine" where the independent participants reach agreement not only about transactional values, 
but also about their order - naturally forming the underlying structure of what is now known as a blockchain. 
"Proof-of-work is essentially one-CPU-one-vote" and this is the novelty 
introduced by Bitcoin \cite{pass2016hybrid,pass2016hybrid2}. The focus shifted 
for decentralized systems and after proof-of-work many other consensus mechanisms have been proposed, 
based on different consensus units, like proof-of-stake, proof-of-space, proof-of-authority, etc. The chaotic 
diversity of new consensus protocols gave also room for endless reviews, overviews and comparisons \cite{survey-dist-consensus, BAMAKAN2020113385, 8400278, 8629877, 9376868, BOURAGA2021114384}. 
The authors of these surveys often put multiple dimensions into comparison, like fault tolerance, scalability, or energy consumption,
and, among those, some focused their efforts on mechanisms that may work in resource-constrained networks \cite{9104713, 8661654, 8168250, SALIMITARI2020100212}. 
This paper will try to identify the common conclusions from these comparisons, while looking into the state-of-the-art and novel approaches for running permissionless
consensus protocols in mobile and edge devices.

@@

There are a couple of particularities that need to be mentioned. First, the block generation
is of probabilistic nature, as the nonce is a random value, and the hash function is a one-way function.
Allied to this, increasing the hashing power is not an easy task, which consequently allows for tackling Sybil attacks,
despite the permissionless and pseudonymity nature of the network. Plus, the incentive mechanism
also plays a role in this regard, by naturally incentivizing honest participation.
The block generation interval, forged by the adjustment of the difficulty
threshold, is also of major importance because it reduces the fork probability and
allows for the timely propagation of proposed blocks, guaranteeing consensus
in the presence of an honest majority of the hashing power \cite{garay2015bitcoin}.
This probabilistic finality shifts the 1/3 BFT threshold to a 1/2 threshold.

@@

Despite the harmony of all the aforementioned clever characteristics,
the proof-of-work mechanism has some well known drawbacks. 
First, the block generation is computationally expensive, and so, the
energy consumption associated with it is also high. This is caused by the
dynamically adjusted difficulty threshold, which is a function of the block 
generation interval. For example, in the Bitcoin network, to maintain the 10 
minutes interval, the hashing difficulty
increases with the increase of the hashing power, and so too the energy needed for computing that proof. 
However, reducing the block generation interval would increase the
probability of forks, which results in a more frequent waste of useful work,
loosening the security of the network. Increasing the block size would also
have the same security effect due to the increased propagation delay \cite{8629877}.

As pointed in the literature, the design of these consensus mechanisms
shall aim for a protocolar choice between a set of properties that form a trilemma (see Fig. \ref{fig:trilemma}): 
Security, Scalability, and Decentralization.
Summarizing, relaxing the security requirements may allow for more scalability, both of which,
consequently, have hands tied with decentralization. These trade-offs are of practical
consideration when defining the network goals and use cases \cite{survey-dist-consensus}.

@@

Idealized and inspired by Proof-of-Stake, extending or adapting Proof-of-Work
became a popular trend in the blockchain community. The main idea is to
replace the computational power with some other resource, that is
more scarce, or more valuable, or more verifiable, or more traceable, etc.,
or to combine multiple resources, or even to add extra requirements to 
pure Proof-of-Work. From Proof-of-Meaningful-Work, or Proof-of-Luck, 
to Proof-of-Work-Time, going through Proof-of-Humanity, or Proof-of-Personhood, there
are countless and exotic variations of these mechanisms, and the list is growing \cite{token-economy-gitbook}.
Not that every one of them has a big potential for entirely solving the permissionless consensus problem,
but each one of them may tackle different use cases where consensus needs to be reached, 
and where different resources are available to guarantee and make the agreement happen \cite{BOURAGA2021114384}.

@@

\subsubsection{Proof-of-Elapsed-Time} is another alternative to Proof-of-Work,
especially targeting the intensive mining issue, that has been proposed 
by Intel. The idea is to simulate the mining process by using a
randomized time delay, which is proportional to the amount of work
that would be required to mine the block. The time delay is then used to pace the
block generation and proposal, and the block acceptance can be performed
using a BFT protocol \cite{olson2018sawtooth}. This solution, however,
is not a viable alternative to other, permissioned and permissionless, approaches, 
by holding strong and centralized hardware requirements, 
as the computations require a trusted execution environment
(TEE) to establish the trust among the participants. Some efforts have been made
to combine the TEE feature with Proof-of-Stake, in order to achieve a
more decentralized and secure protocol, but the 
same hardware requirements are still a major issue.

@@

\subsubsection{DAG structures} may diverge from the previous approaches with an envisioned 
non-linear structure for the result of the consensus attestations. The choice resides in 
directed acyclic graphs that should expand and physically connect, further in time, the new
data appended to the network. The data, which is typically held in the vertices of the graph, 
can have more or less granularity, either by storing blocks of transactions, or raw transactions directly.
This design decision may pose new potential challenges, related to the verification and finalization of
the consensus protocol. Currently, those two approaches are then distinguished as \emph{blockDAG} and 
\emph{txDAG} \cite{survey-dist-consensus}. 
The block-based DAG approach diverges from the traditional blockchain structure by flexibly 
connecting every new block to multiple previous blocks. The rest of the protocol may work in the same way 
as the traditional Proof-of-Work, however, this new feature yields new security considerations. These protocols
are theoretically designed for very high throughput in permissionless environments, but the security trade-offs
are still not well-known \cite{sompolinsky2018phantom}.
In a transaction-based DAG, every vertex holds one transaction, that is connected to previous vertices. 
All the sibling vertices shall then represent disjoint transactions to successfully deal with history and conflict
resolutions. Two well-known examples of this approach are IOTA Tangle \cite{popov2018tangle} 
and Nano \cite{lemahieu2018nano}.
The former is of particular interest, as it is especially designed for IoT applications, low-powered devices, and
micro-transactions. The protocol states that every new vertex needs to be connected to two previous vertices, 
therefore approving the parent transactions, as an incentive mechanism for honest behavior \cite{survey-dist-consensus}. 
The users are the ones to submit transactions, generating and attaching a PoW proof to it and 
so guaranteeing probabilistic finality with 1/2 fault tolerance. This solution may indeed pair well with 
decentralization and scalability, however, there are known security issues related to the
easiness of growing malicious parallel branches with little computational effort, and those issues have been 
addressed by the introduction of an undesirable centralized solution. 
Other interesting and worth mentioning protocols are the ones that combine the DAG structure with the BFT consensus, 
such as the Snowflake-Avalanche protocol \cite{rocket2018snowflake}.

>>>>

\usepackage[pdftex]{graphicx}
\graphicspath{ {./images/} }
\usepackage{caption}

@@

The advancements of the internet 
more than potentiated the revolution and what we now call the permissionless consensus problem was 
finally born. Without forgetting the previous attempts, the first practical permissionless consensus algorithm was 
proposed by Nakamoto in \cite{nakamoto2008bitcoin}. It is a proof-of-work consensus protocol that resembles a "replicated 
state machine" where the independent participants reach agreement not only about transactional values, 
but also about their order - naturally forming the underlying structure of what is now known as a blockchain. 
"Proof-of-work is essentially one-CPU-one-vote" and this is the novelty 
introduced by Bitcoin \cite{pass2016hybrid,pass2016hybrid2}. The focus shifted 
for decentralized systems and after proof-of-work many other consensus mechanisms have been proposed, 
based on different consensus units, like proof-of-stake, proof-of-space, proof-of-authority, etc. The chaotic 
diversity of new consensus protocols gave also room for endless reviews, overviews and comparisons \cite{survey-dist-consensus, BAMAKAN2020113385, 8400278, natoli2019deconstructing, 8629877, 9376868, BOURAGA2021114384}. 
The authors of these surveys often put multiple dimensions into comparison, like fault tolerance, scalability, or energy consumption,
and, among those, some focused their efforts on mechanisms that may work in resource-constrained networks \cite{9104713, 8661654, 8168250, SALIMITARI2020100212}. 
This paper will try to identify the common conclusions from these comparisons, while looking into the state-of-the-art and novel approaches for running permissionless
consensus protocols in mobile and edge devices.

@@

There are a couple of particularities that need to be mentioned. First, the block generation
is of probabilistic nature, as the nonce is a random value, and the hash function is a one-way function.
Allied to this, increasing the hashing power is not an easy task, which consequently allows for tackling Sybil attacks,
despite the permissionless and pseudonymity nature of the network. Plus, the incentive mechanism
also plays a role in this regard, by naturally incentivizing honest participation.
The block generation interval, forged by the adjustment of the difficulty
threshold, is also of major importance because it reduces the fork probability and
allows for the timely propagation of proposed blocks, guaranteeing consensus
in the presence of an honest majority of the hashing power \cite{garay2015bitcoin, natoli2019deconstructing}.
This probabilistic finality shifts the 1/3 BFT threshold to a 1/2 threshold.

@@

Despite the harmony of all the aforementioned clever characteristics,
the proof-of-work mechanism has some well known drawbacks. 
First, the block generation is computationally expensive, and so, the
energy consumption associated with it is also high. This is caused by the
dynamically adjusted difficulty threshold, which is a function of the block 
generation interval. For example, in the Bitcoin network, to maintain the 10 
minutes interval, the hashing difficulty
increases with the increase of the hashing power, and so too the energy needed for computing that proof. 
However, reducing the block generation interval would increase the
probability of forks, which results in a more frequent waste of useful work,
loosening the security of the network. Increasing the block size would also
have the same security effect due to the increased propagation delay \cite{8629877, natoli2019deconstructing}.

As pointed in the literature, the design of these consensus mechanisms
shall aim for a protocolar choice between a set of properties that form a trilemma (See Fig. \ref{fig:trilemma}): 
Security, Scalability, and Decentralization.
Summarizing, relaxing the security requirements may allow for more scalability, both of which,
consequently, have hands tied with decentralization. These trade-offs are of practical
consideration when defining the network goals and use cases \cite{survey-dist-consensus}.

@@

Idealized and inspired by Proof-of-Stake, extending or adapting Proof-of-Work
became a popular trend in the blockchain community. The main idea is to
replace the computational power with some other resource, that is
more scarce, or more valuable, or more verifiable, or more traceable, etc.,
or to combine multiple resources, or even to add extra requirements to 
pure Proof-of-Work. From Proof-of-Meaningful-Work, or Proof-of-Luck, 
to Proof-of-Work-Time, going through Proof-of-Humanity, or Proof-of-Personhood, there
are countless and exotic variations of these mechanisms, and the list is growing \cite{token-economy-gitbook, 9376868}.
Not that every one of them has a big potential for entirely solving the permissionless consensus problem,
but each one of them may tackle different use cases where consensus needs to be reached, 
and where different resources are available to make the agreement happen \cite{BOURAGA2021114384, 9376868}.

@@

\subsubsection{Proof-of-Elapsed-Time} is another alternative to \emph{PoW},
especially targeting the intensive mining issue, that has been proposed 
by Intel. The idea is to simulate the mining process by using a
randomized time delay, which is proportional to the amount of work
that would be required to mine the block. The time delay is then used to pace the
block generation and proposal, and the block acceptance can be performed
using a BFT protocol \cite{olson2018sawtooth}. This solution, however,
is not a viable alternative to other, permissioned and permissionless, approaches, 
by holding strong and centralized hardware requirements, 
as the computations require a trusted execution environment
(TEE) to establish the trust among the participants. Some efforts have been made
to combine the TEE feature with Proof-of-Stake, in order to achieve a
more decentralized and secure protocol, but the 
same hardware requirements are still a major issue.

\subsubsection{Proof-of-Location} may be an alternative for envisioning a future where the
consensus is achieved by taking advantage of the dynamism and easiness of movement, placement and 
tracking of smaller and more location-versatile devices, such as mobile phones, IoT devices, etc.
The idea behind this approach is to use the location and the movement of the devices 
as a proof for the consensus. It relies on the location validation by close proximity, 
for example, using low-range communication channels, either by proving the location
of the device at a certain point in time, or by proving their proximity to other devices.
Incentive mechanisms may play an important role to prevent location spoofing, collusion, etc. \cite{natoli2019deconstructing}.
This protocol is getting increased attention, especially in the context of IoT, 
with the provision of decentralized, permissionless and autonomous networks of 
low-powered devices that can be used for providing these location verification services \cite{9376868}.

\begin{figure*}[h]
  \centering
  \captionsetup{justification=centering}
  \includegraphics[width=\textwidth]{proof-alternative-consensus}
  \caption{From \cite{9376868}, an extensive list of pure proof alternative consensus mechanisms.}
  \label{fig:proof-alternative-consensus}
\end{figure*}

@@

\subsubsection{DAG structures} may diverge from the previous approaches with an envisioned 
non-linear structure for the result of the consensus attestations. The choice resides in 
directed acyclic graphs that should expand and physically connect, further in time, the new
data appended to the network. The data, which is typically held in the vertices of the graph, 
can have more or less granularity, either by storing blocks of transactions, or raw transactions directly.
This design decision may pose new potential challenges, related to the verification and finalization of
the consensus protocol. Currently, those two approaches are then distinguished as \emph{blockDAG} and 
\emph{txDAG} \cite{survey-dist-consensus}. 
The block-based DAG approach diverges from the traditional blockchain structure by flexibly 
connecting every new block to multiple previous blocks. The rest of the protocol may work in the same way 
as the traditional Proof-of-Work, however, this new feature yields new security considerations. These protocols
are theoretically designed for very high throughput in permissionless environments, but the security trade-offs
are still not well-known \cite{sompolinsky2018phantom}.
In a transaction-based DAG, every vertex holds one transaction, that is connected to previous vertices. 
All the sibling vertices shall then represent disjoint transactions to successfully deal with history and conflict
resolutions. Two well-known examples of this approach are IOTA Tangle \cite{popov2018tangle} 
and Nano \cite{lemahieu2018nano}.
The former is of particular interest, as it is especially designed for IoT applications, low-powered devices, and
micro-transactions. The protocol states that every new vertex needs to be connected to two previous vertices, 
therefore approving the parent transactions, as an incentive mechanism for honest behavior \cite{survey-dist-consensus}. 
The users are the ones to submit transactions, generating and attaching a \emph{PoW} proof to it and 
so guaranteeing probabilistic finality with 1/2 fault tolerance. This solution may indeed pair well with 
decentralization and scalability, however, there are known security issues related to the
easiness of growing malicious parallel branches with little computational effort, and those issues have been 
addressed by the introduction of an undesirable centralized solution. 
Other interesting and worth mentioning protocols are the ones that combine the DAG structure with the BFT consensus, 
such as the Snowflake-Avalanche protocol \cite{rocket2018snowflake}.
